---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Alina Maldonado Kellig (12900411), Sara Rainer (12969031), Lars Mehrtens (11953802)"
date: "`r Sys.Date()`"
output: distill::distill_article
---

## Exercise 1: Initializing git 
    
### d)

2 strengths:
Great for collaborations, which is especially useful for work within companies,
and if you’ve committed properly, you can always go back to any point within the project and revert the steps to see where a possible mistake happened.

2 weaknesses:
Git is very complex, which makes it difficult for beginners to learn.
Git also struggles with large files so it might become really slow when you try to store really huge files.
    

## Exercise 2: Putting your Repository on GitHub 

### a)
    
    https://github.com/Sarao7/exeRcise-sheet-3.git 
    

## Exercise 3: Pixar Films 

```{r}
# Create a directory for raw data
dir.create("data/raw", recursive = TRUE)

# Download raw datasets
download.file(
  url = "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/pixar_films.csv",
  destfile = "data/raw/pixar_films_raw.csv"
)

download.file(
  url = "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/public_response.csv",
  destfile = "data/raw/public_response_raw.csv"
)
```

### a)
    
```{r}
library(tidyverse)

# Load raw data
pixar_films <- read_csv("data/raw/pixar_films_raw.csv")

# Clean data - remove films without titles
pixar_clean <- pixar_films %>%
  filter(!is.na(film))
```
```{r}
# Inspect film_rating variable
unique(pixar_clean$film_rating)
```

Film rating meaning: Rating based on Motion Picture Association (MPA) film rating system (the dataset only includes "G" and "PG"):

  1. G - General Audiences
  2. PG - Parental Guidance Suggested
  3. PG-13 - Parents Strongly Cautioned
  4. R - Restricted

```{r}
# Create factor variable for film rating
pixar_clean <- pixar_clean %>%
  mutate(film_rating_factor = factor(film_rating,
                                   levels = c("G", "PG", "PG-13", "R"),
                                   ordered = TRUE))

```


<!-- --> Why factor is appropriate:

  1. Film ratings are categorical with inherent ordering (G < PG < PG-13 < R)
  2. Factors preserve this ordinal relationship for analysis and visualization
  3. Prevents treating ratings as numeric values
  4. Ensures proper handling in statistical models
    
### b)

Toy Story, Cars and the Incredibles form a series (Toy Story friom part 1 to 4, Cars from part 1 to 3, the Incredibles from part 1 to 2). Technically also Finding Nemo/ Dory and Monsters Inc./ University are series.

```{r}
library(stringr)

# create dataframe that lists how many films are part of a series
pixar_series <- pixar_clean %>%
  mutate(
    series_name = case_when(
      str_detect(film, "Toy Story") ~ "Toy Story",
      str_detect(film, "Cars") ~ "Cars",
      str_detect(film, "Finding") ~ "Finding Nemo/Dory",
      str_detect(film, "Incredibles") ~ "The Incredibles",
      str_detect(film, "Monsters") ~ "Monsters Inc.",
      str_detect(film, "A Bug's Life") ~ "A Bug's Life",
      TRUE ~ film 
    )
  ) %>%
  group_by(series_name) %>%
  summarize(
    films = paste(film, collapse = ", "),
    films_count = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(films_count), series_name)

print(pixar_series)
```

    
### c)

```{r}
# Load the public_response dataset
public_response <- read_csv("data/raw/public_response_raw.csv")

# Convert cinema_score to ordered factor
public_response <- public_response %>%
  mutate(
    cinema_score = factor(
      cinema_score,
      levels = c("F", "D", "C", "B", "A", "A+"),
      ordered = TRUE
    )
  )

# Verify the factor levels
levels(public_response$cinema_score)


# Merge with pixar_films dataset with film column
combined_data <- pixar_clean %>%
  inner_join(public_response, 
             by = "film")

```

    
### d)

```{r}
# put series variable in the combined data
combined_with_series <- combined_data %>%
  mutate(
    series = case_when(
      str_detect(film, "Toy Story") ~ "Toy Story",
      str_detect(film, "Cars") ~ "Cars",
      str_detect(film, "Finding") ~ "Finding Nemo/Dory",
      str_detect(film, "Incredibles") ~ "The Incredibles",
      str_detect(film, "Monsters") ~ "Monsters Inc.",
      TRUE ~ "Standalone"
    )
  ) %>%
  filter(series != "Standalone") %>%
  group_by(series) %>%
  mutate(series_size = n()) %>%
  filter(series_size > 1) %>%  # Only keep actual series
  arrange(release_date) %>%   # Order by release date
  ungroup()

# create score plot with metacritic variable
ggplot(combined_with_series,
       aes(x = series,
           y = metacritic,
           fill = fct_inorder(film))) +  # Order by release date
  geom_col(position = position_dodge(width = 0.8),
           width = 0.7) +
  labs(
    title = "Metacritic Scores Across Pixar Film Series",
    subtitle = "Ordered by release date within each series",
    x = "Film Series",
    y = "Metacritic Score (0-100)",
    fill = "Film Title"
  ) +
  theme_minimal()
```


Observation: Usually the first part of a series is rated higher than the follow-ups.

## Exercise 4: Open Analysis


### b)

```{r}
# Define the URL of the raw CSV file
url <- "https://raw.githubusercontent.com/owid/owid-datasets/master/datasets/Global%20CO2%20emissions%20-%20CDIAC%20and%20UN%20Population/Global%20CO2%20emissions%20-%20CDIAC%20and%20UN%20Population.csv"

# Define the local path to save the file
destfile <- "Global CO2 emissions - CDIAC and UN Population.csv"

# Download the file
download.file(url, destfile)

```


### c)

-   Dataset Name: Global CO₂ Emissions – CDIAC and UN Population
-   Link: https://github.com/owid/owid-datasets/tree/master/datasets/Global%20CO2%20emissions%20-%20CDIAC%20and%20UN%20Population

    
### d)  

Research Question: How have global per capita CO₂ emissions changed from 1850 to 2015, and can we identify any long-term trends?

Why this is interesting: It reveals whether individuals globally are becoming more or less carbon-intensive over time — a critical question in climate change mitigation and policy planning.

    
### e)
    
```{r}
# Read data
co2 <- read_csv("Global CO2 emissions - CDIAC and UN Population.csv")

# Rename for simplicity
co2_clean <- co2 %>%
  rename(
    year = Year,
    total_co2 = `CO2 emissions (CDIAC and UN Population)`,
    per_capita_co2 = `Per capita CO2 emissions (CDIAC and UN Population)`
  ) %>%
  filter(Entity == "World", year >= 1850, year <= 2015)

# Inspect
head(co2_clean)

```
```{r}
ggplot(
  co2_clean, 
  aes(
    x = year, 
    y = per_capita_co2
    )
  ) +
  geom_line(color = "steelblue", size = 1) +
  labs(
    title = "Global Per Capita CO2 Emissions (1850–2015)",
    x = "Year",
    y = "Tonnes CO2 per Person",
    caption = "Figure 1. Trends in global per capita CO2 emissions based on CDIAC and UN population data (Our World in Data)."
  ) +
  theme_minimal()

```


Interpretation:
Per capita CO2 emissions have steadily increased from 1850 to about the 1970s.

The fluctuations from 1920 - 1950 might be because of the reduced industrial production during the Great Depression and the destruction of means of production during World War 2.

There's a plateau and slight decline post-2007, reflecting changes in energy efficiency, economic shifts, or climate policy impacts.

The graph shows clear historical inflection points during global industrialization and energy transitions.

How have global per capita CO₂ emissions changed from 1850 to 2020?
Yes, we can see CO2 emissions clearly increasing dramatically in this period.

Can we identify any long-term trends?
Eventhough CO2 emissions have increased globally since 1850, you can see a small indication of a trend in the different direction. Through renewable energy and degrowth politics this trend doesnt necessarily have to continue.


### f)
    
```{r co2_plot, fig.cap="Figure: Global CO₂ emissions and per capita emissions from 1850 to 2015."}
#duplicate dataset
co2_f <- co2

# Rename columns
names(co2_f) <- c("Entity", "Year", "Total_CO2", "Per_Capita_CO2")

# Reshape data to long format for ggplot
co2_long <- pivot_longer(co2_f, 
                         cols = c(Total_CO2, Per_Capita_CO2),
                        names_to = "Type", 
                        values_to = "Value")

# Plot
ggplot(
  co2_long, 
  aes(
    x = Year, 
    y = Value, 
    color = Type
    )
  ) +
  geom_line(size = 1.2) +
  labs(
    title = "Global Total vs. Per Capita CO2 Emissions (1850–2015)",
       x = "Year",
       y = "CO2 Emissions",
       color = "Metric") +
  scale_color_manual(
    labels = c("Per Capita CO2 (tonnes)", "Total CO2 (billion tonnes)"),
    values = c("Per_Capita_CO2" = "steelblue", "Total_CO2" = "firebrick")) +
  theme_minimal()

```
