---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Alina Maldonado Kellig (12900411), Sara Rainer (12969031), Lars Mehrtens (11953802)"
date: "`r Sys.Date()`"
output: distill::distill_article
---

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `2025-exeRcise-sheet-3`. If you forgot how to do this, you can
    follow this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)
    
### d)

2 strengths:
Great for collaborations, which is especially useful for work within companies,
and if you’ve committed properly, you can always go back to any point within the project and revert the steps to see where a possible mistake happened.

2 weaknesses:
Git is very complex, which makes it difficult for beginners to learn.
Git also struggles with large files so it might become really slow when you try to store really huge files.
    
e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves. You
    can, but don't need to create a `.gitignore` file. Just do not
    commit these files manually.

## Exercise 2: Putting your Repository on GitHub (3 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:
    
    https://github.com/Sarao7/exeRcise-sheet-3.git 
    
b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.

## Exercise 3: Pixar Films (4 Points)

Download the `pixar_films` and `public_response` datasets from the
GitHub repository and track them in git.

Link:
<https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-03-11>

For small datasets like these adding them to git is not a problem.

```{r}
# Create a directory for raw data
dir.create("data/raw", recursive = TRUE)

# Download raw datasets
download.file(
  url = "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/pixar_films.csv",
  destfile = "data/raw/pixar_films_raw.csv"
)

download.file(
  url = "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/public_response.csv",
  destfile = "data/raw/public_response_raw.csv"
)
```


a)  Load the `pixar_films` dataset into R. Clean the dataset by removing
    films without a title. Inspect the variable `film_rating`. What are
    the possible values and what do they mean? Create a factor variable
    for the film rating. Why is this appropriate?
    
### a)
    
```{r}
library(tidyverse)

# Load raw data
pixar_films <- read_csv("data/raw/pixar_films_raw.csv")

# Clean data - remove films without titles
pixar_clean <- pixar_films %>%
  filter(!is.na(film))
```
```{r}
# Inspect film_rating variable
unique(pixar_clean$film_rating)
```

Film rating meaning: Rating based on Motion Picture Association (MPA) film rating system (the dataset only includes "G" and "PG"):

  1. G - General Audiences
  2. PG - Parental Guidance Suggested
  3. PG-13 - Parents Strongly Cautioned
  4. R - Restricted

```{r}
# Create factor variable for film rating
pixar_clean <- pixar_clean %>%
  mutate(film_rating_factor = factor(film_rating,
                                   levels = c("G", "PG", "PG-13", "R"),
                                   ordered = TRUE))

```


<!-- --> Why factor is appropriate:

  1. Film ratings are categorical with inherent ordering (G < PG < PG-13 < R)
  2. Factors preserve this ordinal relationship for analysis and visualization
  3. Prevents treating ratings as numeric values
  4. Ensures proper handling in statistical models

b)  Inspect the film titles manually. Which films form a film series? A
    film series can be identified by a common word in the titles of the
    films, often in conjunction with a number in the title,
    e.g. "Despicable Me" and "Despicable Me 2". Create a dataframe which
    displays a list of the different series with the titles of the films
    and how many films belong to the series. Output the dataframe.
    
### b)

Toy Story, Cars and the Incredibles form a series (Toy Story friom part 1 to 4, Cars from part 1 to 3, the Incredibles from part 1 to 2). Technically also Finding Nemo/ Dory and Monsters Inc./ University are series.

```{r}
library(stringr)

# create dataframe that lists how many films are part of a series
pixar_series <- pixar_clean %>%
  mutate(
    series_name = case_when(
      str_detect(film, "Toy Story") ~ "Toy Story",
      str_detect(film, "Cars") ~ "Cars",
      str_detect(film, "Finding") ~ "Finding Nemo/Dory",
      str_detect(film, "Incredibles") ~ "The Incredibles",
      str_detect(film, "Monsters") ~ "Monsters Inc.",
      str_detect(film, "A Bug's Life") ~ "A Bug's Life",
      TRUE ~ film 
    )
  ) %>%
  group_by(series_name) %>%
  summarize(
    films = paste(film, collapse = ", "),
    films_count = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(films_count), series_name)

print(pixar_series)
```


c)  Load the `public_response` dataframe into R. Convert the
    `cinema_score` variable into a factor while ensuring the factor
    levels are defined in ascending order, from the lowest to the
    highest score. Combine `public_response` with the `pixar_films`
    dataset using an appropriate merge variable.
    
### c)

```{r}
# Load the public_response dataset
public_response <- read_csv("data/raw/public_response_raw.csv")

# Convert cinema_score to ordered factor
public_response <- public_response %>%
  mutate(
    cinema_score = factor(
      cinema_score,
      levels = c("F", "D", "C", "B", "A", "A+"),
      ordered = TRUE
    )
  )

# Verify the factor levels
levels(public_response$cinema_score)


# Merge with pixar_films dataset with film column
combined_data <- pixar_clean %>%
  inner_join(public_response, 
             by = "film")

```


d)  Choose one of the variables representing the public response and
    create a bar plot for the films belonging to a series. Here are the
    details of the plot:

    -   The film series are represented on the x-axis.
    -   Your chosen public response variable is displayed on the y-axis.
    -   Each film in the series is represented as a separate bar. Bars
        are grouped by film under their respective series on the x-axis.
        Order the bars within a series according to the release date of
        the films.
    -   A title and axis labels for context.

    What do you notice when comparing the scores of the films in a
    series? Do you see any patterns?
    
### d)

```{r}
# put series variable in the combined data
combined_with_series <- combined_data %>%
  mutate(
    series = case_when(
      str_detect(film, "Toy Story") ~ "Toy Story",
      str_detect(film, "Cars") ~ "Cars",
      str_detect(film, "Finding") ~ "Finding Nemo/Dory",
      str_detect(film, "Incredibles") ~ "The Incredibles",
      str_detect(film, "Monsters") ~ "Monsters Inc.",
      TRUE ~ "Standalone"
    )
  ) %>%
  filter(series != "Standalone") %>%
  group_by(series) %>%
  mutate(series_size = n()) %>%
  filter(series_size > 1) %>%  # Only keep actual series
  arrange(release_date) %>%   # Order by release date
  ungroup()

# create score plot with metacritic variable
ggplot(combined_with_series,
       aes(x = series,
           y = metacritic,
           fill = fct_inorder(film))) +  # Order by release date
  geom_col(position = position_dodge(width = 0.8),
           width = 0.7) +
  labs(
    title = "Metacritic Scores Across Pixar Film Series",
    subtitle = "Ordered by release date within each series",
    x = "Film Series",
    y = "Metacritic Score (0-100)",
    fill = "Film Title"
  ) +
  theme_minimal()
```


Observation: Usually the first part of a series is rated higher than the follow-ups.

## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from
[Our World in Data](https://ourworldindata.org/) and analyze it, while
determining the research question yourself.

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets>
    and choose a dataset that interests you. You can have a look at
    <https://ourworldindata.org/> to gather some inspiration.
    
b)  Download the dataset and track it in git.

### b)

```{r}
# Define the URL of the raw CSV file
url <- "https://raw.githubusercontent.com/owid/owid-datasets/master/datasets/Global%20CO2%20emissions%20-%20CDIAC%20and%20UN%20Population/Global%20CO2%20emissions%20-%20CDIAC%20and%20UN%20Population.csv"

# Define the local path to save the file
destfile <- "Global CO2 emissions - CDIAC and UN Population.csv"

# Download the file
download.file(url, destfile)

```

c)  Put the name / title of the dataset and a link to it below.

### c)

-   Dataset Name: Global CO₂ Emissions – CDIAC and UN Population
-   Link: https://github.com/owid/owid-datasets/tree/master/datasets/Global%20CO2%20emissions%20-%20CDIAC%20and%20UN%20Population

d)  Come up with a (research) question you want to answer with the data
    and briefly explain why you believe this is an interesting question
    within one sentence. It should be a question that can be answered
    with the dataset and using R.
    
### d)  

Research Question: How have global per capita CO₂ emissions changed from 1950 to 2020, and can we identify any long-term trends?

Why this is interesting: It reveals whether individuals globally are becoming more or less carbon-intensive over time — a critical question in climate change mitigation and policy planning.

e)  Use R to answer your chosen question. Please limit your analysis to
    the functions and techniques we have covered so far in the course.
    You are **not expected** to use advanced statistical models or
    external packages which haven't been introduced.
f)  Create a meaningful plot / figure with the dataset. Make sure to
    provide a figure caption (via the chunk options / Rmarkdown) and
    correctly label the figure.

## Final Note

Make sure to push all your commits and changes to GitHub before
submitting the exercise sheet.
